---
title: "Assignment 02-QMM"
author: "Chandima Attanayake"
date: "`r Sys.Date()`"
output:
  word_document: default
  html_document:
    df_print: paged
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
**The Problem has been solved under the following basis.**

# 1. Objective

Maximize the profit = (420 X y no of Large) + (360 X y no of Medium) + (300 X y no of small)

# 2. Constraints and Assumptions 

**Capacity constraints (per plant) - Each plant has a maximum capacity**

•	Plant 1: yLarge +yMedium + ySmall ≤ 750

•	Plant 2: yLarge +yMedium + ySmall ≤ 900

•	Plant 3: yLarge +yMedium + ySmall ≤ 450

**Storage constraints - Storage space per plant**

•	20 × yLarge +15 × yMedium + 12 × ySmall ≤ 13,000 for Plant 01

•	20 × yLarge +15 × yMedium + 12 × ySmall ≤ 12,000 for Plant 02

•	20 × yLarge +15 × yMedium + 12 × ySmall ≤ 5,000 for Plant 03

**Sales constraints - The maximum number of each size that can be sold**

•	yLarge plant 1+ yLarge plant 2 + yLarge plant 3 ≤ 900 for large

•	yMedium plant 1+ yMedium plant 2 + yMedium plant 3 ≤ 1,200 for medium 

•	ySmall plant 1+ ySmall plant 2 + ySmall plant 3 ≤ 750 for small

**Equal capacity usage: Each plant uses the same percentage of its total capacity**


```{r}
# Load the lpSolve library
library(lpSolve)

# Coefficients of the objective function (profit per unit)
objective <- c(420, 360, 300, 420, 360, 300, 420, 360, 300)  # Large, Medium, Small for 3 plants

# Constraints matrix
constraints <- matrix(c(
  # Capacity constraints
  1, 1, 1, 0, 0, 0, 0, 0, 0,  # Plant 1 (<= 750)
  0, 0, 0, 1, 1, 1, 0, 0, 0,  # Plant 2 (<= 900)
  0, 0, 0, 0, 0, 0, 1, 1, 1,  # Plant 3 (<= 450)
  
  # Storage constraints
  20, 15, 12, 0, 0, 0, 0, 0, 0,  # Plant 1 (<= 13000 sq ft)
  0, 0, 0, 20, 15, 12, 0, 0, 0,  # Plant 2 (<= 12000 sq ft)
  0, 0, 0, 0, 0, 0, 20, 15, 12,  # Plant 3 (<= 5000 sq ft)
  
  # Sales constraints
  1, 0, 0, 1, 0, 0, 1, 0, 0,  # Large sizes (<= 900)
  0, 1, 0, 0, 1, 0, 0, 1, 0,  # Medium sizes (<= 1200)
  0, 0, 1, 0, 0, 1, 0, 0, 1   # Small sizes (<= 750)
), nrow = 9, byrow = TRUE)

# RHS of the constraints
rhs <- c(750, 900, 450, 13000, 12000, 5000, 900, 1200, 750)

# Constraint direction
direction <- c("<=", "<=", "<=", "<=", "<=", "<=", "<=", "<=", "<=")

# Solve the linear programming problem
solution <- lp("max", objective, constraints, direction, rhs)

# Display the results
solution$objval  # The maximum profit
solution$solution  # The optimal production plan

```